Designing Large-Scale Imaging Surveys for a Retrospective Relative Photometric Calibration

### 2012-01-16 paper/self_calibration_I.tex
euclid: fix title error in paper
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20089 03a3cea6-2c03-0410-ac9b-9271b1c66c29
euclid: paper post HW comments
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@20142 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\shorttitle{Designing Large Scale Imaging Surveys for Retrospective Relative Photometric Calibration}
\shortauthors{Holmes et al.}

@@

\title{Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}

@@

\begin{abstract}
In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate the large-scale response of imaging instruments used in large-scale imaging surveys with survey science data alone. We create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a synthetic sky according to a defined survey strategy. These catalog-based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response. We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey dataset, based on the multiple observations of (non-varying) sources at different focal plane positions. By considering four simple survey strategies we find that, with appropriate redundancy built into the survey strategy, it is possible to accurately constrain the relative photometric response of an imaging instrument, and therefore the relative calibration of all the images and the relative fluxes of all the sources observed. The majority of the remaining post-self-calibration errors are due to the limitations in the basis used to model the relative instrument response. We find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate calibration. The results of this work suggest the following advice for those considering the design of large scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
\end{abstract}

@@

\section{Introduction}
Astronomers tend to think in terms of taking science data and calibration data separately. The former is used for scientific measurements of astronomical sources and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on. But typically far more photons, or readout electrons, are collected during science exposures themselves; are these not incredibly constraining on the calibration? Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}. But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: Good redundancy is required in the data stream, and a redundancy of a very specific kind.

@@

We construct realistic survey datasets through mock observations of a synthetic sky according to a specified survey strategy. These simulations include a complex, position-dependent instrument response for the imaging instrument. Through the self-calibration procedure, we recover this instrument response by fitting a model that best describes the survey dataset. By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies. There is a degeneracy in this problem: the self-calibration procedure is only able to fit for a \textit{relative} instrument response. We have no way to know, for example, if all the sources are uniformly fainter, or if the instrument sensitivity is uniformly lower.

In this work we do not produced pixelated images; instead we concentrate on catalog-level simulations with realistic measurement uncertainties. Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the dataset.

In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey datasets. Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure, with Section \ref{sec:metrics} summarizing the metrics we use to assess its performance. In Section \ref{sec:simple_surveys} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties.

\section{Survey Dataset Simulations}
\label{sec:survey_simulations}
We have constructed an end-to-end simulation chain that produces a realistic imaging dataset from a specified survey strategy. In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated. The simulations are split into a number of steps. The first is the generation of a synthetic sky. With a given pointing, single mock observations are performed on this sky. To build up a full-survey dataset, multiple single observations are performed according to the specified survey strategy. In this section, we detail the assumptions and methods used in each of these steps.

\subsection{The Synthetic Sky}
We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and $10\upsigma{}$ limits of deep, space-based, near-infrared survey. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
\begin{eqnarray}
\log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
\end{eqnarray}
where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters. Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citet{win11} only. These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$. The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy). To limit the data, we only select the brightest sources within the survey area, up to a source density per unit solid angle $d$, for the self-calibration procedure.

\subsection{A Single Exposure}
\label{sec:single_exposure}
With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$ this synthetic sky is transformed into focal plane image. In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers. An example of a single pointing exposure is shown in Figure \ref{fig:single_image}.

\subsubsection{Measured Count Rates}
We do not consider pixelated images in these simulations; instead the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$ 
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response, and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

\subsubsection{Noise Model}
\label{sec:noise}
To construct the noise model, the simulated exposures are assumed to be background limited and that, for systematic reasons, there is an upper limit on the signal-to-noise ratio of 500 for bright sources. The noise model is complicated further by applying an extra term $\epsilon_i$ to the count rates' uncertainty variance, which we intentionally do not take into account in the analysis in order to simulate systematic problems with the instrument noise model. The \textit{true} noise model is therefore
\begin{eqnarray}
\sigma_{i, \true}^{2} = (1 + \epsilon_i) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} ]^2 \quad , \label{eqn:noise}
\end{eqnarray}
where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, drawn uniformly in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the 500 limit on the signal-to-noise ratio are used to set $\alpha = 0.1585$ and $\eta = 0.0017$. The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are \textit{assumed} (incorrectly, but realistically) to be
\begin{eqnarray*}
\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2} \, c^{2}_i \quad 
\end{eqnarray*}
during this stage.

\subsubsection{The True Instrument Response Model}
\label{sec:instrument_response_model}
We construct a complex, position independent instrument response model $f_\true(\vec{x_i} | \vec{q}_{\true})$ from a superposition of large and small-scale variations:
\begin{eqnarray*}
f_\true(\vec{x_i} | \vec{q}_{\true, 1 \ldots 260}) = f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6}) + f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7 \ldots 260}) \quad ,
\end{eqnarray*}
where $\vec{x_i} = (x_i, y_i)$ is the focal plane position that the $k$th source falls at during the $i^\text{th}$ measurement and $\vec{q}_\true$ are the parameters defining the instrument response model. The large-scale instrument response $f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6})$ is modeled as a second order polynomial:
\begin{eqnarray*}
f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6}) = q_{\true, 1} + q_{\true, 2} \, x_i + q_{\true, 3} \, y_i + q_{\true, 4} \, x_i^2 + q_{\true, 5} \, x_i \, y_i  + q_{\true, 6} \, y_i^2  \quad .
\end{eqnarray*}
The small-scale instrument response, which is constructed from sine and cosine contributions, is superimposed on this large-scale instrument response. The small-scale instrument response $f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7 \ldots 260})$ is modeled as
\begin{eqnarray*}
f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
& & \qquad \qquad \times \left[ q_{\true, 9+4b} \cos (k_y \, y_i) + q_{\true, 10+4b} \sin (k_y \, y_i) \right] \quad ,
\end{eqnarray*}
where
\begin{eqnarray*}
k_x & = & \dfrac{a \, \pi}{X} \quad ,\\
k_y & = & \dfrac{b \, \pi}{Y} \quad ,\\
\end{eqnarray*}
with the physical focal plane dimensions $X$ and $Y$. In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image}      (right). It is this instrument response that we try and recover with the self-calibration procedure. Our final assumption in these simulations is that the instrument response is temporally stable.

\subsection{The Complete Survey}
In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting dataset. We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy. The resultant source measurements are collated into a survey-wide dataset.

\section{Self-Calibrating the Survey-Wide Dataset}
\label{sec:self_cal}
We self-calibrate the dataset generated in the survey simulations to recover a best possible estimate of the true instrument response applied and the true source fluxes. This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''} and the Deep Lens Survey \citep{wit11}. This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates. These steps are iterated until the system converges, or until it is clear that the system will not converge. There is a degeneracy in the problem, as both the true instrument response and the true source magnitudes are unknown. It is therefore only possible to calibrate the \textit{relative} instrument response and the \textit{relative} source fluxes. It is not possible to know, for example, if the sources are all fainter or if the instrument response is uniformly lower. In practice, this degeneracy can be broken through observations of a small number of standard absolute sources.

\subsection{Fitted Measurement Model}
To complicate the simulations, we use the self-calibration procedure to fit a model that is \textit{incomplete} in two ways. Firstly, the \textit{fitted} instrument response is modeled as an eighth order polynomial, and not the second order polynomial superimposed with sine and cosine contributions used to model the \textit{true} instrument response. Secondly, the assumed measurement uncertainty variances do not include the additional random measurement error $\epsilon_{i}$ introduced in Subsection \ref{sec:noise}. The incomplete measurement model is
\begin{eqnarray*}
c_i = f(\vec{x_i} | \vec{q}) \, s_{k} + e_{i} \quad ,
\end{eqnarray*}
where $c_i$ is the recorded count rate, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal plane position $\vec{x}_i$, $\vec{q}$ is a vector parameterizing the eighth order polynomial instrument response model, $s_k$ is the model source flux estimate and the error $e_i$ is drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, such that
\begin{eqnarray*}
\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2}\, c^{2}_i \quad ,
\end{eqnarray*} 
where $\alpha$ and $\eta$ are the parameters set by the instruments $10\upsigma$ and saturation limits. The $\epsilon_i$ error contribution is intentionally not included in order to simulate systematic problems with the instrument noise model.

We make one further assumption during this analysis phase: We assume that clearly variable sources have been removed from the source catalog, and that the remaining sources vary by less than $\epsilon_{max}$.

@@

\subsection{Step 2: Instrument Response Refinement}
The instrument response parameters can now be refined with the latest source flux estimates. A error function for all the measurements of all the sources is constructed
\begin{eqnarray*}
\chi^2 = \sum_{k} \chi'^2_k \quad ,
\end{eqnarray*}
where
\begin{eqnarray*}
\chi'^2_k = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})^2}{{\sigma_i}^2}  \quad .
\end{eqnarray*}
Recall that the fitted instrument response $f_{i}(\vec{x_i} | \vec{q})$ is modeled as an eighth order polynomial. This can be expressed as
\begin{eqnarray*}
f_{i}(\vec{x_i} | \vec{q}) = \sum_{l = 1}^L q_{l} \, g_l(\vec{x_i})  \quad ,
\end{eqnarray*}
where $L = 45$ in this case. The total error function $\chi^2$ can be rewritten as
\begin{eqnarray*}
\chi^2 =\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{l} g_l(\vec{x_i}))^2}{{\sigma_i}^2}   \quad .
\end{eqnarray*}
To refine the instrument response model fit, this error function is minimized with respect to the instrument response model parameters $q_{l}$
\begin{eqnarray*}
\frac{d\chi^2}{dq_{l}} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i}^2} = 0  \quad ,
\end{eqnarray*}
\begin{eqnarray*}
\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i}^2} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'} (\vec{x_i})} {{\sigma_i}^2}   \quad .
\end{eqnarray*}
It is now simpler to proceed in matrix notation. The following substitutions can be made
\begin{eqnarray}
b_l & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{\sigma_i^2}    \quad ,\\
G_{ll'} & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{{s'}_{k}^2}{\sigma_i^2} g_l(x_i) g_{l'}(x_i)   \quad .
\end{eqnarray}
The matrix equation is then
\begin{eqnarray*}
\vec{b} = G \cdot \vec{q'}    \quad .
\end{eqnarray*}
The refined instrument response parameters are then found by
\begin{eqnarray*}
\vec{q'} \leftarrow G^{-1}  \cdot \vec{b}   \quad .
\end{eqnarray*}
These two steps are iterated until the solution converges to a final fit of the instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ and the source fluxes $s_{k,\fit}$, or until it is clear that a solution will not be found.

@@

\section{Simple Survey Strategies}
\label{sec:simple_surveys}
In this paper, we consider four simple but very different survey strategies. We use the end-to-end dataset simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies.

>>>>

\shorttitle{Designing Large-Scale Imaging Surveys for Retrospective Relative Photometric Calibration}
\shortauthors{Holmes et al.}

@@

\title{Designing Large-Scale Imaging Surveys for a Retrospective Relative Photometric Calibration}

@@

\begin{abstract}
In this paper we show that, with a well chosen observing strategy, it is possible to accurately calibrate the large-scale response of instruments used in large-scale imaging surveys with survey science data alone. We create end-to-end simulations of an imaging survey, which produce simulated photometric catalogs from mock observations of a synthetic sky according to a defined survey strategies. These catalog-based simulations include realistic measurement uncertainties and a complex, position-dependent instrument response. We then use a self-calibration technique to recover the relative instrument response by fitting an model that best explains the survey catalog, based on (non-varying) sources that fall on different focal plane positions in multiple observations. By considering four simple survey strategies we find that, with appropriate redundancy built into the survey strategy, it is possible to accurately constrain the large-scale relative photometric response of an imaging instrument, and therefore the relative calibration of all the images and the relative fluxes of all the sources observed. The majority of the remaining post-self-calibration errors are due to the limitations in the basis used to model the relative instrument response. We find that returning the same sources to very different focal plane positions is the key property of any survey strategy designed for accurate retrospective calibration. The results of this work suggest the following advice for those considering the design of large-scale imaging surveys: Do not use a regular tiling of the sky; instead return the same sources to very different focal plane positions.
\end{abstract}

@@

\section{Introduction}
Astronomers tend to think conventionally in terms of taking science data and calibration data separately. The former is used for scientific measurements of astronomical sources and the latter is used to constrain instrument parameters, such as the instrument response, the dark currents and so-on. But typically far more photons, or readout electrons, are collected during science exposures themselves; are these not incredibly constraining on the calibration? Indeed, in the retrospective photometric calibration of the Sloan Digital Sky Survey data (SDSS), much more calibration information was obtained from the science data than the calibration data \citep{pad08}. But, of course, the SDSS imaging strategy had to be adjusted to make this calibration work: Good redundancy was required in the data stream, and a redundancy of a very specific kind.

@@

We construct realistic survey catalogs through mock observations of a synthetic sky according to a specified survey strategy. These simulations include a complex, position-dependent instrument response for the imaging instrument. Through the self-calibration procedure, we recover this instrument response by fitting a model that best describes the survey catalog. By comparing the \textit{fitted} instrument response to the \textit{true} instrument response we are able to assess the performance of the self-calibration procedure with different survey strategies.

In this work we do not produced pixelated images; instead we concentrate on catalog-level simulations with realistic measurement uncertainties. Complex effects are included in the simulations that are not  precisely modeled at the analysis stage, in order to simulate the effects of unknown systematic errors within the catalog.

In Section \ref{sec:survey_simulations} we introduce the simulation chain constructed to produce the realistic survey catalogs. Section \ref{sec:self_cal} goes onto to detail the self-calibration procedure, with Section \ref{sec:metrics} summarizing the metrics we use to assess its performance. In Section \ref{sec:simple_surveys} we focus on four simple survey strategies, which allow us to draw conclusions on the performance of the self-calibration procedure with different survey properties.

\section{Survey catalog Simulations}
\label{sec:survey_simulations}
We have constructed an end-to-end simulation chain that produces a realistic imaging catalog from a specified survey strategy. In this work we have kept the simulation parameters intentionally flexible, so that the sensitivity of the self-calibration procedure to different values can be investigated (the fiducial values are given in Table \ref{tab:parameters}). The simulations are split into a number of steps. The first is the generation of a synthetic sky. With a given pointing, single mock observations are performed on this sky. To build up a full-survey catalog, multiple single observations are performed according to the specified survey strategy. In this section, we detail the assumptions and methods used in each of these steps.

\subsection{The Synthetic Sky}
We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and a $10\upsigma{}$ limits of deep, space-based, near-infrared survey. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
\begin{eqnarray}
\log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
\end{eqnarray}
where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters. Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citet{win11} only. These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$. The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy). To limit the data, we only select the brightest sources within the survey area, up to a source density per unit solid angle $d$, for the self-calibration procedure.

\subsection{A Single Exposure}
\label{sec:single_exposure}
With a camera pointing $(\alpha, \beta)$ and camera orientation $\theta$, we find the focal plane positions of the sources on the synthetic sky that fall within the instrument's field-of-view. In our simulations, we use a large instrument field-of-view of $0.76~ \text{deg} \times 0.72~\text{deg}$; a size consistent with the next generation of large survey imagers. An example of a single pointing exposure is shown in Figure \ref{fig:single_image}.

\subsubsection{Measured Count Rates}
To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$ the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s_{k,\true}$ ,
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response (see Section \ref{sec:instrument_response_model}), and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

\subsubsection{Noise Model}
\label{sec:noise}
To construct the noise model, the simulated exposures are assumed to be background limited and that, for systematic reasons, there is an upper limit on the signal-to-noise ratio of 500 for bright sources. The noise model is complicated further by applying an extra term $\epsilon_i$ to the count rates' uncertainty variance, which we intentionally do not take into account in the analysis in order to simulate systematic problems with the instrument noise model. The \textit{true} noise model is therefore
\begin{eqnarray}
\sigma_{i, \true}^{2} = (1 + \epsilon_i) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} ]^2 \quad , \label{eqn:noise}
\end{eqnarray}
where $\alpha$ and $\eta$ are both constants and $\epsilon_i$ is a random number, drawn uniformly in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. The $m = 22$ mag 10$\upsigma$ detection limit introduced previously and the limit S/N$<500$ are used to set $\alpha = 0.1585$ and $\eta = 0.0017$. The $\epsilon_i$ contribution is not taken into account in the analysis and therefore the uncertainty variances on the count rates are \textit{assumed} (incorrectly, but realistically) to be
\begin{eqnarray*}
\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2} \, c^{2}_i \quad 
\end{eqnarray*}
during this stage.

\subsubsection{The True Instrument Response Model}
\label{sec:instrument_response_model}
We construct a complex, position independent instrument response model $f_\true(\vec{x_i} | \vec{q}_{\true})$ from a superposition of large and small-scale variations:
\begin{eqnarray*}
f_\true(\vec{x_i} | \vec{q}_{\true, 1 \ldots 260}) = f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6}) + f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7 \ldots 260}) \quad ,
\end{eqnarray*}
where $\vec{x_i} = (x_i, y_i)$ is the focal plane position that the $k$th source falls at during the $i\text{th}$ measurement and $\vec{q}_\true$ are the parameters defining the instrument response model. This response model subsumes various effects commonly discussed in terms of detector flat-fielding and optical vignetting. The large-scale instrument response $f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6})$ is modeled as a second order polynomial:
\begin{eqnarray*}
f_\text{large}(\vec{x_i} | \vec{q}_{\true, 1 \ldots 6}) = q_{\true, 1} + q_{\true, 2} \, x_i + q_{\true, 3} \, y_i + q_{\true, 4} \, x_i^2 + q_{\true, 5} \, x_i \, y_i  + q_{\true, 6} \, y_i^2  \quad .
\end{eqnarray*}
The small-scale instrument response, which is constructed from sine and cosine contributions, is superimposed on this large-scale instrument response. The small-scale instrument response $f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7 \ldots 260})$ is modeled as
\begin{eqnarray*}
f_\text{small}(\vec{x_i} | \vec{q}_{\true, 7\ldots 260})  & = &  \sum_{a=0}^6 \sum_{b=0}^a \left[ q_{\true, 7+4b} \cos (k_x \, x_i) + q_{\true, 8+4b} \sin (k_x \, x_i) \right] \\
& & \qquad \qquad \times \left[ q_{\true, 9+4b} \cos (k_y \, y_i) + q_{\true, 10+4b} \sin (k_y \, y_i) \right] \quad ,
\end{eqnarray*}
where
\begin{eqnarray*}
k_x & = & \dfrac{a \, \pi}{X} \quad ,\\
k_y & = & \dfrac{b \, \pi}{Y} \quad ,\\
\end{eqnarray*}
with the physical focal plane dimensions $X$ and $Y$. In total, the instrument response model is parameterized with 260 parameters; an example can be seen in Figure \ref{fig:single_image}      (right). It is this instrument response that we try and recover with the self-calibration procedure. With this instrument response model, we have assumed that the variations in sensitivities on a pixel-to-pixel-scale have been calibrated with some other method, such as an internal flat-field calibration source. It is variations on larger scales that we aim to constrain with the self-calibration procedure. Our final assumption in these simulations is that the instrument response is temporally stable.

\subsection{The Complete Survey}
In this work we are interested in simulating complete surveys, in order to identify the crucial characteristics of survey strategies that allow for the relative instrument response model to be accurately, retrospectively constrained from the resulting catalog. We therefore apply the single exposure procedure introduced in the previous section for each pointing specified in a defined survey strategy. The resultant source measurements are collated into a survey-wide catalog.

\section{Self-Calibrating the Survey-Wide catalog}
\label{sec:self_cal}
We self-calibrate the catalog generated in the survey simulations to recover an optimum estimate of the true instrument response applied and the true source fluxes. This self-calibration procedure has been successfully applied to ground-based imaging surveys, such as the Sloan Digital Sky Survey \citep{pad08}\footnote{Here this procedure is dubbed ``uber-calibration.''} and the Deep Lens Survey \citep{wit11}. This iterative procedure comprises two steps: (1) a refinement of the source flux estimates based on the latest instrument response model and (2) a refinement of the instrument response model based on the updated source flux estimates. These steps are iterated until the system converges, or until it is clear that the system will not converge. There is a degeneracy in the problem, as both the true instrument response and the true source magnitudes are unknown. It is therefore only possible to calibrate the \textit{relative} instrument response and the \textit{relative} source fluxes. It is not possible to know, for example, if the sources are all fainter or if the instrument response is uniformly lower. In practice, this degeneracy can be broken through observations of a small number of standard absolute sources.

\subsection{Fitting the Instrument Response Model}
To introduce a realistic level of ignorance into the simulations, we use the self-calibration procedure to fit a model that is \textit{incomplete} in two ways. Firstly, the \textit{fitted} instrument response is modeled as an eighth order polynomial, and not the second order polynomial superimposed with sine and cosine contributions used to model the \textit{true} instrument response. Secondly, the assumed measurement uncertainty variances do not include the additional random measurement error $\epsilon_{i}$ introduced in Subsection \ref{sec:noise}. We 
We therefore fit for the following, incomplete measurement model:
\begin{eqnarray*}
c_i = f(\vec{x_i} | \vec{q}) \, s_{k} + e_{i} \quad ,
\end{eqnarray*}
where $c_i$ is the recorded count rate, $f(\vec{x_i} | \vec{q})$ is the fitted instrument response model at a focal plane position $\vec{x}_i$, $\vec{q}$ is a vector parameterizing the eighth order polynomial instrument response model, $s_k$ is the model source flux estimate and the error $e_i$ is drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, such that
\begin{eqnarray*}
\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2}\, c^{2}_i \quad ,
\end{eqnarray*} 
where $\alpha$ and $\eta$ are the parameters set by the instrument's $10\upsigma$ detection and the S/N $<500$ limits. The $\epsilon_i$ error contribution is intentionally not included in order to simulate systematic problems with the instrument noise model.

We make one further assumption during this analysis phase, namely that clearly variable sources have been removed from the source catalog, and that the remaining sources vary by less than $\epsilon_{max}$.

@@

\subsection{Step 2: Instrument Response Refinement}
The instrument response parameters can now be refined with the latest source flux estimates. A error function for all the measurements of all the sources is constructed
\begin{eqnarray*}
\chi^2 = \sum_{k} \chi'^2_k \quad ,
\end{eqnarray*}
where
\begin{eqnarray*}
\chi'^2_k = \sum_{i \in \mathcal{O}(k)} \frac{(c_i-f_{i}(\vec{x_i} | \vec{q}) \, s'_{k})^2}{{\sigma_i'}^2}  \quad .
\end{eqnarray*}
Recall that the fitted instrument response $f_{i}(\vec{x_i} | \vec{q})$ is modeled as an eighth order polynomial. This can be expressed as
\begin{eqnarray*}
f_{i}(\vec{x_i} | \vec{q}) = \sum_{l = 1}^L q_{l} \, g_l(\vec{x_i})  \quad ,
\end{eqnarray*}
where $L = 45$ in this case. The total error function $\chi^2$ can be rewritten as
\begin{eqnarray*}
\chi^2 =\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{(c_i- s'_{k} \sum_{l = 1}^L q_{l} g_l(\vec{x_i}))^2}{{\sigma_i'}^2}   \quad .
\end{eqnarray*}
To refine the instrument response model fit, this error function is minimized with respect to the instrument response model parameters $q_{l}$
\begin{eqnarray*}
\frac{d\chi^2}{dq_{l}} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{-2 g_l(\vec{x_i}) s'_{k} (c_i- s'_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'}(\vec{x_i}))}{{\sigma_i'}^2} = 0  \quad ,
\end{eqnarray*}
\begin{eqnarray*}
\sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{{\sigma_i'}^2} = \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l (\vec{x_i}) s'^2_{k} \sum_{l' = 1}^{L'} q'_{l'} g_{l'} (\vec{x_i})} {{\sigma_i'}^2}   \quad .
\end{eqnarray*}
It is now simpler to proceed in matrix notation. The following substitutions can be made
\begin{eqnarray}
b_l & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{g_l(\vec{x_i}) s'_{k} c_i}{\sigma_i'^2}    \quad ,\\
G_{ll'} & = & \sum_{k} \sum_{i \in \mathcal{O}(k)} \frac{{s'}_{k}^2}{\sigma_i'^2} g_l(x_i) g_{l'}(x_i)   \quad .
\end{eqnarray}
The matrix equation is then
\begin{eqnarray*}
\vec{b} = G \cdot \vec{q'}    \quad .
\end{eqnarray*}
The refined instrument response parameters are then found by
\begin{eqnarray*}
\vec{q'} = G^{-1}  \cdot \vec{b}   \quad .
\end{eqnarray*}
These two steps are iterated until the solution converges to a final fit of the instrument response $f_\fit(\vec{x} | \vec{q_\fit})$ and the source fluxes $s_{k,\fit}$, or until it is clear that a solution will not be found.

@@

\section{Simple Survey Strategies}
\label{sec:simple_surveys}
In this paper, we consider four simple but very different survey strategies. We use the end-to-end catalog simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies.

\subsection{Survey Description}
We label the four survey strategies -- which all cover the same patch of sky -- with the letters A to D. These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}. Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~\% and $\sim 8$~\% in the camera pointing directions $\alpha$ and $\beta$ respectively). The pointings in the 12 passes over the same field are exactly aligned. Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $30^\circ$. Survey C is more complex. The first pass over the field is the same as in Survey A, with $12 \times 12$ pointings. In the next pass, one of the pointings in the $\alpha$ direction is removed and one is added in the $\beta$ and the pointings are respaced uniformly, so the resultant pointing grid is $13 \times 11$. In the third pass over the field this change is reversed and the field is measured on a $11 \times 13$ grid. These three passes are then repeated four times. The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.35,0.35)~deg applied in both the $\alpha$ and $\beta$ directions. By fixing the pointings within these $0.7~\text{deg} \times 0.7~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field. The orientations of the pointings in Survey D are completely random.
